A Generalization of Regularized Discriminant Analysis for High-Dimensional Classification

### 2013-04-13 reports/paper-rda.Rnw
Added and polished Caleb's writeup of simulation results. Also, reformatted the paragraphs through the paper using the magnificent power of emacs.

In this section, we describe four high-dimensional microarray data sets and compare our proposed \emph{GRDA} classifier with five recently proposed classifiers designed for small-sample, high-dimensional data: the \emph{DLDA} classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA} (\emph{SCRDA}) classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} (\emph{SDLDA}) classifier from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA (\emph{SmDLDA}) classifier from \cite{Tong:2012hw}, and the penalized \emph{LDA} (\emph{PLDA}) classifier from \cite{Witten:2011kc}. We did not include the \emph{RDA} classifier from \cite{Friedman:1989tm} in our studies to avoid its excessive computations.  We evaluated the classification performance of each classifier by randomly partitioning the data such that $2/3$ of the observations were allocated as training data and the remaining $1/3$ of the observations were allocated as a test data set. For each random partition, we applied the feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features having the largest ratio of their between-group to within-group sums of squares from the training data. We then filtered the test data set with the corresponding $d$ features. After training each classifier from the filtered training data, we classified the filtered test data sets and computed the proportion of mislabeled test observations to estimate the classification error rate for each classifier. Repeating this process 1000 times, we computed the average of the error-rate estimates for each classifier. We remark that our simulation design avoids the selection bias discussed in \cite{Ambroise:2002fa}.  In Table \ref{tab:data-summary}, we summarize the attributes of each microarray data set.

We used version 2.15.1 of the open-source statistical software {\tt R} for the classification study. For each classifier's {\tt R} implementation, we used the default settings with the exception that we explicitly set the prior probabilities as equal. In our {\tt R} implementation of the \emph{GRDA} classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1, \ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$ from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold cross-validation \citep{Hastie:2008dt}.

@@

For the microarray data sets, we computed error-rate estimates for the competing classifiers after reducing the original feature space dimension with variable selection to $d=100,250,500$, and $1000$.  Our choices for the values of $d$ are similar to those of \cite{Xu:2009fl}.  We report the error-rate estimates along with approximate standard errors in parentheses in Tables \ref{tab:results-d100} -- \ref{tab:results-d1000} along with the chosen reduced dimension $d$.

The \emph{GRDA} classifier was superior in classification performance for the majority of the simulations. While the error-rate estimates for the majority of the classifiers increased as $d$ was increased, the classification accuracy for the \emph{GRDA} classifier improved as $d$ increased for the Burczynski and Nakayama data sets. For the Nakayama data set, the \emph{GRDA} classifier was outperformed by the Dudoit, Pang, Tong, and Witten classifiers for small values of $d$ but attained classification accuracy comparable to its competitors for larger values of $d$. Moreover, the only situation where the other classifiers considered were competitive with the \emph{GRDA} classifier was in the case of the Nakayama data set.

The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were comparable regardless of the choice of $d$. In particular, the error-rate estimates for Dudoit's and Tong's methods were nearly identical, which suggests that the improved mean estimator employed by Tong provides little improvement to the classification performance in the \emph{DLDA} classifier. The popular classifier from Guo is competitive only for large values of $d$ for the Shipp and Singh data sets. For the Burczynski and Nakayama data sets, the implementation of Guo's method in the {\tt rda} package often classified all of the test observations into a single class, thus yielding larger error rates than its competitors.

>>>>

In this section, we describe four high-dimensional microarray data sets and
compare our proposed \emph{GRDA} classifier with five recently proposed
classifiers designed for small-sample, high-dimensional data: the \emph{DLDA}
classifier from \cite{Dudoit:2002ev}, the shrunken centroids \emph{RDA}
classifier from \cite{Guo:2007te}, the shrinkage-based \emph{DLDA} classifier
from \cite{Pang:2009ik}, the shrinkage-mean-based DLDA classifier from
\cite{Tong:2012hw}, and the penalized \emph{LDA} classifier from
\cite{Witten:2011kc}. Below, we refer to each classifier by the first author's
surname. We evaluated the classification performance of each classifier by
randomly partitioning the data such that $2/3$ of the observations were
allocated as training data and the remaining $1/3$ of the observations were
allocated as a test data set. For each random partition, we applied the
feature-selection method from \cite{Dudoit:2002ev} to choose the $d$ features
having the largest ratio of their between-group to within-group sums of squares
from the training data. We then filtered the test data set with the
corresponding $d$ features. After training each classifier from the filtered
training data, we classified the filtered test data sets and computed the
proportion of mislabeled test observations to estimate the classification error
rate for each classifier. Repeating this process 1000 times, we computed the
average of the error-rate estimates for each classifier. We remark that our
simulation design avoids the selection bias discussed in
\cite{Ambroise:2002fa}. We did not include the \emph{RDA} classifier from
\cite{Friedman:1989tm} in our studies to avoid its excessive computations. In
Table \ref{tab:data-summary}, we summarize the attributes of each microarray
data set.

We used version 3.0.0 of the open-source statistical software {\tt R} for the
classification study. For each classifier's {\tt R} implementation, we used the
default settings with the exception that we explicitly set the prior
probabilities as equal. In our {\tt R} implementation of the \emph{GRDA}
classifier in the {\tt regdiscrim} package, we set $\alpha_k = 1$, $k = 1,
\ldots, K$ for simplicity. We estimated $\lambda$ from a grid of 21 equidistant
candidate values between 0 and 1, inclusively. Similarly, we estimated $\gamma$
from a grid consisting of the values $10^{-1}, \ldots, 10^4$, and $10^5$. We
selected optimal estimates of $\lambda$ and $\gamma$ using 10-fold
cross-validation \citep{Hastie:2008dt}.

@@

For the microarray data sets, we computed error-rate estimates for the competing
classifiers after reducing the dimension of the data sets with feature selection
to $d = 100, \ldots, 1000$, similar to the values of $d$ considered by
\cite{Xu:2009fl}. To compare misclassification rates between different values of
$d$, we applied each value of $d$ considered to the same random allocation of
training and test data. Hence, for a given random partition, the set of features
selected for a small value of $d$ is a proper subset of the features selected
for a larger value of $d$. We report the error-rate estimates along with
approximate standard errors in parentheses in Tables \ref{tab:results-d100} --
\ref{tab:results-d1000} along with the reduced dimension $d$.

The \emph{GRDA} classifier was superior in classification performance for the
majority of the simulations. For the Burczysnki, Shipp, and Singh data sets, the
\emph{GRDA} classifier outperformed its competitors for all values of $d$
examined. For the Nakayama data set, the \emph{GRDA} classifier was outperformed
by the Dudoit, Pang, and Tong classifiers for small values of $d$ but attained
classification rates superior to its competitors for larger values of
$d$. Furthermore, Figure \textbf{TODO} emphasizes two strengths of the
\emph{GRDA} classifier: reliability of the classifier for small $d$ and
stability of the classifier for large $d$. In contrast to the other classifiers,
the \emph{GRDA} classifier exhibited less variability its misclassification
estimates for small values of $d$. Also, the \emph{GRDA} classifier yielded
superior classification accuracy as $d$ increased and was hardly effected by the
inclusion of more variables in the data sets. We attribute the apparent
stability of the error-rate estimates of the \emph{GRDA} classifier for large
values of $d$ to its novel dimension-reduction technique.

The error-rate estimates for the Dudoit, Pang, Tong, and Witten classifiers were
comparable for all values of $d$ examined. Furthermore, the error-rate estimates
of the Dudoit, Pang, and Tong classifiers were approximately equal throughout
our study, and the Witten classifier yielded approximately equal error-rate
estimates for the Burczysnki, Shipp, and Singh data sets. Furthermore, the
error-rate estimates for the Dudoit and Tong classifiers were nearly identical,
which suggests that the improved mean estimator employed by \cite{Tong:2012hw}
provided little improvement to the classification performance of the \emph{DLDA}
classifier.

The classification performance of the the Dudoit, Pang, Tong, and Witten
classifiers for the Shipp and Singh data sets revealed the inability of these
methods to handle larger values of $d$. After closer examination of the
classification studies performed in the authors' original papers, we noticed
that classification accuracy estimates were reported only for small values of
$d$, thereby ignoring the inclusion of a larger number of variables. Hence, we
advise that practioners be aware of the necessity to filter a large number of
features before employing these methods, lest their classification performance
can degrade.

The popular classifier from Guo is most competitive over large values of $d$ for
the Shipp and Singh data sets. However, for the Burczynski and Nakayama data
sets, the implementation of Guo's method in the {\tt rda} package often
classified every test observation into a single class, thus yielding larger
error rates than its competitors. Furthermore, the Guo classifier exhibited
unreliable classification performance for smaller values of $d$. The Guo
classifier required a larger subset of the features to perform competitively in
our study.

As seen in Figure ?, the \emph{GRDA} classifier exhibits an interesting bathtub
shape in selection of the lambda tuning parameter through cross-validation. (Not
sure what else to write here.)
